---
title: "Getting started with apsimR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with apsimR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!-- Render time: ~13 s under rmarkdown::render() with APSIM Next Generation
     installed (dominated by one real simulator call); a build with no APSIM
     installed renders in well under a second, taking the abstention path
     throughout. macOS arm64, R 4.5.2, one core. -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4.5,
  dpi = 150
)
```

```{r library}
library(apsimR)
has_apsim <- apsimR::apsim_available()
```

## Why

An agronomist who already has an APSIM Next Generation file open asks one
question before anything else: can I drive this simulator from R, get back a
result with enough provenance that a causal or decision tool downstream can
trust it, and will my analysis still tell me something useful, rather than
simply fail, on a laptop where APSIM itself is not installed? This vignette
answers all three parts of that question by walking every verb apsimR
exposes: reading and running a simulation, fitting a reduced-form mechanism
to its output, building the shared forward-model closure that calibration,
sensitivity analysis and emulation all reduce to, validating simulated
output against observations, generating a known-causal-structure dataset to
grade a causal method against, and handing every result on into the
orchestra's shared manifest contract.

## What

`apsim_sim()` reads an `.apsimx` file into a typed object that round-trips
through `apsim_write()`; `apsim_available()`, `apsim_version()` and
`apsim_configure()` report and control what simulator apsimR finds, and
every verb that needs the simulator returns a typed `apsim_abstention`
(checked with `is_apsim_abstention()`) rather than an error when it is
absent. `apsim_predict()` runs the simulator once and returns its report as
a `predictions` manifest; `apsim_estimate()` runs it across a varying input
and fits an interpretable dose-response to the result. Calibration,
sensitivity analysis and emulation are the same operation seen three ways --
vary parameters, run the simulator, reduce each run to an observation
vector -- so all three are built over one shared closure,
`apsim_forward_model()`, fed by parameter points from `apsim_design()`:
`apsim_calibrate()` solves the inverse problem (which parameters reproduce
observed data), `apsim_sensitivity()` ranks which parameters drive an output,
and `apsim_emulate()` fits a fast Gaussian-process surrogate for one.
`apsim_validate()` scores simulated output against observations with typed
goodness-of-fit metrics, and `apsim_ground_truth()` uses the simulator's
ability to reveal both potential outcomes of the same unit to manufacture a
dataset with a known causal answer, for grading a causal method that could
never see that answer from a real experiment. Every verb returns an
`apsim_manifest`; `as_orchestra_manifest()` and `as_pesto_manifest()` carry
that result into the wider orchestra.

## Do

### Is the simulator available, and what happens when it is not?

apsimR is standalone-functional: it installs and loads with no simulator
present, and every simulator-dependent verb abstains, returning a typed
reason, rather than erroring, when APSIM is absent.

```{r availability}
apsim_available()
apsim_version()
```

When the simulator cannot be found, a verb returns an `apsim_abstention` a
caller can branch on. The demonstration below deliberately points apsimR at
a non-existent `Models` executable to force that path, then reverts to
auto-discovery.

```{r abstention}
# A minimal simulation file, and a deliberately wrong simulator path.
f <- tempfile(fileext = ".apsimx")
writeLines(paste0('{"$type":"Models.Core.Simulations, Models",',
                  '"Name":"Simulations","Children":[]}'), f)
apsim_configure(models = file.path(tempdir(), "no-such-Models"))

res <- apsim_predict(f)
is_apsim_abstention(res)
res$reason

apsim_configure() # revert to auto-discovery
```

### Reading a simulation file

An `.apsimx` file is a typed tree of model nodes. `apsim_sim()` reads it; the
object round-trips through `apsim_write()`. Reading needs no simulator run,
but the object records the installed APSIM version, so this section still
depends on `has_apsim`.

```{r sim, eval = has_apsim, echo = has_apsim}
sim <- apsim_sim(apsim_example("Wheat"))
sim
apsim_simulations(sim)
```

```{r sim-skip, eval = !has_apsim, echo = FALSE, results = "asis"}
cat("APSIM Next Generation is not available on this build, so the shipped",
    "Wheat example is not read here; `apsim_sim()` itself needs no",
    "simulator run, but the section above shows the abstention every",
    "simulator-dependent verb falls back to instead.\n")
```

### Prediction

`apsim_predict()` runs the simulator and returns the simulated reports as a
`predictions` manifest -- a contract-shaped object carrying the outputs plus
provenance (the APSIM version, an integrity hash).

```{r predict, eval = has_apsim, echo = has_apsim}
pred_m <- apsim_predict(apsim_example("Wheat"))
pred_m
```

```{r predict-skip, eval = !has_apsim, echo = FALSE, results = "asis"}
cat("APSIM Next Generation is not available on this build, so no",
    "prediction is run here; `apsim_predict()` would return the same",
    "typed abstention demonstrated above.\n")
```

### Estimation: fitting a reduced-form mechanism

`apsim_estimate()` runs the simulator across an input and fits a reduced-form
mechanism (a saturating Mitscherlich dose-response by default), returning the
fitted parameters as a `parameters` manifest -- the interpretable summary a
causal or decision member consumes. The default form is the classical
saturating N-response curve attributed to Mitscherlich (1909),

$$y(x) = y_0 + y_{\max}\left(1 - e^{-r x}\right),$$

with asymptote $y_{\max}$, zero-input intercept $y_0$ and curvature $r$
governing how quickly the response saturates (the Limits section below notes
how this attribution was checked). `apsim_estimate()` fits this form by
nonlinear least squares through the internal helper it calls,
`apsimR:::.apsim_fit_response()`, which is what runs underneath every
`apsim_estimate(sim, x, y)` call once the simulator is available. The
demonstration below drives that helper directly on a synthetic N-response,
so it needs no simulator run:

```{r estimate-fit}
set.seed(1)
n_rate <- seq(0, 200, by = 10)
yield <- 1.1 + 4.2 * (1 - exp(-0.018 * n_rate)) +
  stats::rnorm(length(n_rate), 0, 0.1)
est_params <- apsimR:::.apsim_fit_response(n_rate, yield,
                                           response = "mitscherlich")
est_m <- apsim_manifest("parameters", "demo", params = est_params)
```

```{r estimate-kable, echo = FALSE}
knitr::kable(
  est_m@params, digits = 3,
  caption = paste0(
    "Mitscherlich parameters fitted to the synthetic N-response demonstration."
  )
)
```

### The forward model: one machine for three verbs

Calibration, sensitivity analysis and emulation are the same operation seen
three ways -- vary some APSIM parameters, run the simulator, reduce each run
to a small observation vector. `apsim_forward_model()` builds that operation
once, as a closure of the orchestra-standard shape `function(theta) -> obs`:
`theta` is an `nreal x npar` matrix (one row per parameter set), `obs` an
`nreal x nobs` matrix. A single APSIM invocation per row applies the edits
and runs the file in one `--apply` config; failed runs become `NA` rows
rather than aborting the sweep. Building the closure and calling it are two
different costs -- building is free, but calling it runs the simulator once
per row, so the call below is shown for its shape only and is not evaluated
in this render (see the note after the design table for why).

```{r forward, eval = FALSE}
fm <- apsim_forward_model(
  apsim_example("Wheat"),
  parm_paths = "[Fertilise at sowing].Script.Amount",
  output = "Yield")
fm(matrix(c(0, 100, 200), ncol = 1L)) # three nitrogen rates -> three yields
```

`apsim_design()` builds the parameter points a sweep evaluates -- a
space-filling Latin hypercube by default. It is pure (no simulator), so it
runs anywhere:

```{r design}
d_design <- apsim_design(
  c("[Sow].Script.Population", "[Fertilise].Amount"),
  lower = c(50, 0), upper = c(200, 200), n = 5L, seed = 1L)
```

```{r design-kable, echo = FALSE}
knitr::kable(
  d_design, digits = 1,
  caption = "A five-point Latin-hypercube design over two parameters."
)
```

The forward-model call above, and the calibration, sensitivity and full
emulation demonstrations in the next three sections, each invoke the
simulator many times over: a calibration ensemble alone runs it up to
several hundred times. Even on a build where APSIM is installed, running
all of that live inside this vignette would take much longer than its
render-time budget allows, so those calls are shown for their shape only.
Each function's own help page carries a runnable example gated on
`apsim_available()`.

### Calibration

`apsim_calibrate()` solves the inverse problem: find the parameters whose
simulated observations best match data, over a box. The `optim` backend
returns a bounded least-squares point estimate with no extra dependency; the
`ies` backend (default when `PESTO` is installed) returns a posterior
parameter ensemble through an iterative ensemble smoother, with RTPS
inflation guarding against ensemble collapse. Several targets calibrate
jointly through one weighted misfit.

```{r calibrate, eval = FALSE}
apsim_calibrate(
  apsim_example("Wheat"),
  parm_paths = "[Fertilise at sowing].Script.Amount",
  lower = 0, upper = 250,
  observed = c(Yield = 2200), output = "Yield",
  backend = "ies", n_real = 50L, seed = 1L)
```

### Sensitivity analysis

`apsim_sensitivity()` asks which inputs drive an output. Morris screening
(the default) ranks factors by mean absolute elementary effect (`mu_star`)
at a cost of `r(p + 1)` runs and flags interaction or non-linearity
(`sigma`); the Sobol-Jansen method decomposes the output variance into
first-order (`S`) and total (`T`) indices at a higher cost of `n(p + 2)`
runs. Both use the `sensitivity` package over the same forward model.

```{r sensitivity, eval = FALSE}
apsim_sensitivity(
  apsim_example("Wheat"),
  parm_paths = c("[Fertilise at sowing].Script.Amount",
                "[Sow using a variable rule].Script.Population"),
  lower = c(0, 50), upper = c(250, 200),
  output = "Yield", method = "morris", morris_r = 10L, seed = 1L)
```

### Emulation

A single APSIM run costs seconds; a sweep needs thousands. `apsim_emulate()`
runs the simulator at a space-filling design once, fits a Gaussian-process
surrogate, then predicts anywhere in microseconds with a calibrated
uncertainty. Every emulator is returned with mandatory leave-one-out
diagnostics: a surrogate with no measure of its own reliability is worse
than none. The default backend is a genuine exact GP with an
automatic-relevance-determination (ARD) squared-exponential kernel,

$$k(x, x') = \sigma_f^2 \exp\!\left(-\tfrac{1}{2}\sum_{j=1}^{p}
  \frac{(x_j - x_j')^2}{\ell_j^2}\right) + \sigma_n^2 \,\mathbb{1}[x = x'],$$

with a per-dimension length-scale $\ell_j$, signal variance $\sigma_f^2$ and
noise variance $\sigma_n^2$ all fitted by maximising the exact log marginal
likelihood (Rasmussen and Williams, 2006, *Gaussian Processes for Machine
Learning*), not fixed by a heuristic. It needs no extra package and is
exercised directly here (via the internal training/prediction pair
`apsimR:::.apsim_gp_train()` / `apsimR:::.apsim_gp_predict()` that
`apsim_emulate()` itself calls) on a one-dimensional test function, with a
genuine leave-one-out pass -- each point predicted from a GP refit on the
other eleven -- rather than a stand-in smoother:

```{r emulate-demo}
set.seed(1)
truth <- function(x) sin(x) + 0.1 * x
xtr <- matrix(seq(0, 10, length.out = 12), ncol = 1L)
ytr <- truth(xtr[, 1L])

# A genuine leave-one-out pass: refit the GP on the other 11 points and
# predict the held-out one, for each point in turn.
loo <- t(vapply(seq_len(nrow(xtr)), function(i) {
  gp <- apsimR:::.apsim_gp_train(xtr[-i, , drop = FALSE], ytr[-i])
  pr <- apsimR:::.apsim_gp_predict(gp, xtr[i, , drop = FALSE])
  c(mean = pr$mean, sd = pr$sd)
}, numeric(2)))
loo_rmse <- sqrt(mean((ytr - loo[, "mean"])^2))
lo <- loo[, "mean"] - 1.96 * loo[, "sd"]
hi <- loo[, "mean"] + 1.96 * loo[, "sd"]
covered <- ytr >= lo & ytr <= hi
```

```{r fig-loo, echo = requireNamespace("ggplot2", quietly = TRUE), eval = requireNamespace("ggplot2", quietly = TRUE), fig.cap = "Leave-one-out predictions from the exact GP against the true test function, with the fitted 95% interval (shaded) and the GP mean (line). Point colour marks whether the truth fell inside its own held-out interval; the subtitle reports the RMSE and coverage count the caption below also interprets."}
loo_df <- data.frame(
  x = xtr[, 1L], y_true = ytr, y_mean = loo[, "mean"], lo = lo, hi = hi,
  covered = ifelse(covered, "within 95% interval", "outside 95% interval"))
ggplot2::ggplot(loo_df, ggplot2::aes(x = x)) +
  ggplot2::geom_ribbon(ggplot2::aes(ymin = lo, ymax = hi),
                       fill = "grey80", alpha = 0.5) +
  ggplot2::geom_line(ggplot2::aes(y = y_mean), colour = "#0072B2") +
  ggplot2::geom_point(ggplot2::aes(y = y_true, colour = covered), size = 2.6) +
  ggplot2::scale_colour_manual(
    name = NULL,
    values = c("within 95% interval" = "#009E73",
              "outside 95% interval" = "#D55E00")) +
  ggplot2::labs(
    x = "x (arbitrary input units)", y = "y (arbitrary output units)",
    title = "Leave-one-out predictions from the exact GP",
    subtitle = sprintf("RMSE = %.3f | 95%% coverage = %d / %d points",
                       loo_rmse, sum(covered), length(covered))) +
  ggplot2::theme_minimal()
```

```{r fig-loo-skip, echo = FALSE, eval = !requireNamespace("ggplot2", quietly = TRUE), results = "asis"}
cat("ggplot2 is not installed; the leave-one-out figure is skipped.\n")
```

A single twelve-point draw is far too small to certify calibration either
way; the reported `coverage95` from a full `apsim_emulate()` run (typically
evaluated over dozens of held-out points against a design of thirty or more)
is the number to trust, and it can -- and, on at least one independent
Branin-function check, did -- run below its nominal 95%, so read it rather
than assume the interval is calibrated.

On the simulator the same idea fits an emulator of yield against nitrogen,
with `apsim_emulate()` reporting the leave-one-out RMSE, R-squared and 95%
coverage, and `predict()` emulating new rates. Like the calibration and
sensitivity calls above, running this against the real simulator costs
thirty APSIM invocations, so it is shown for its shape only:

```{r emulate, eval = FALSE}
em <- apsim_emulate(
  apsim_example("Wheat"),
  parm_paths = "[Fertilise at sowing].Script.Amount",
  lower = 0, upper = 250, output = "Yield", n = 30L, seed = 1L)
em                                       # prints the LOO diagnostics
predict(em, matrix(c(75, 175), ncol = 1L))
```

### Validation

`apsim_validate()` confronts simulated output with observations and reports
the standard agronomic goodness-of-fit metrics -- error magnitude (RMSE,
MAE), direction (mean error, percent bias) and skill relative to the
observed mean (Nash-Sutcliffe efficiency, Willmott's index of agreement,
RSR) -- with a typed verdict against the performance bands of Moriasi et al.
(2007). It accepts an `apsim_predict()` manifest, a data frame (optionally
joined to the observations on a key such as date), or plain vectors:

```{r validate}
observed <- c(2.1, 3.4, 4.0, 5.2, 6.1)
simulated <- c(2.0, 3.6, 3.8, 5.1, 5.9)
v <- apsim_validate(simulated, observed)
```

```{r validate-kable, echo = FALSE}
knitr::kable(
  v@params, digits = 3,
  caption = paste0(
    "Goodness-of-fit metrics for the five-point validation demonstration",
    " (verdict: ", v@metadata$verdict, ")."
  )
)
```

An observed-versus-predicted plot is available through
`apsim_validation_plot()` when `ggplot2` is installed:

```{r validate-plot, eval = requireNamespace("ggplot2", quietly = TRUE), fig.height = 6, fig.cap = "Simulated against observed values for the five-point demonstration pair, with the 1:1 line of perfect agreement. Points close to the line and on both sides of it (rather than all above or all below) indicate no systematic bias, consistent with the near-zero percent bias reported in the table above."}
apsim_validation_plot(v)
```

```{r validate-plot-skip, echo = FALSE, eval = !requireNamespace("ggplot2", quietly = TRUE), results = "asis"}
cat("ggplot2 is not installed; the observed-versus-predicted figure is",
    "skipped.\n")
```

### Ground truth: APSIM as a causal test-bench

A real experiment never observes both potential outcomes of the same unit,
so the true causal effect is unknowable and a causal method can only be
trusted, never scored. A process simulator escapes that: run the same unit
under control and under treatment and both outcomes are known.
`apsim_ground_truth()` does this across heterogeneous units to manufacture
an observing-system simulation experiment (OSSE) -- a dataset of observed
`(covariates, treatment, outcome)` whose true individual and average
treatment effects are recorded alongside. Treatment can be assigned at
random, or confounded by a covariate that also moves the outcome, so a
causal member (such as the orchestra's TACI / kernR tools) can be handed the
observed data and graded against the truth it cannot see. Each unit needs
two simulator runs (one per potential outcome), so this demonstration is
also shown for its shape only:

```{r ground-truth, eval = FALSE}
units <- data.frame(
  `[Sow using a variable rule].Script.Population` = c(80, 120, 160, 200),
  check.names = FALSE)
gt <- apsim_ground_truth(
  apsim_example("Wheat"),
  treatment = "[Fertilise at sowing].Script.Amount",
  control = 0, treated = 150, units = units, output = "Yield",
  assignment = "confounded",
  confounder = "[Sow using a variable rule].Script.Population", seed = 1L)

gt                   # prints the true ATE and the (biased) naive difference
head(gt@observed)    # (covariates, W, Y) -- the input to a causal method
gt@truth$ate         # the answer key
```

Under confounding the naive treatment-minus-control difference is biased
away from `gt@truth$ate`; a causal method that conditions on the covariates
should recover the recorded ATE. That recovery is the validation gate for
the orchestra's causal members.

### Composing with the orchestra

Every result-returning verb produces an `apsim_manifest`, a versioned,
provenance-complete object of its own. Two adapters carry it into the wider
orchestra.

`as_orchestra_manifest()` emits the federation's general result contract
(`orchestra_manifest`, contract C2 v2 in `ORCHESTRA.md`) at schema
`2.0.0-draft`, and is the route for any apsimR result. A prediction stays a
prediction, a sensitivity table becomes a statement about parameters, and an
emulator or a validation becomes a statement about predictions -- with the
original apsimR target preserved in the manifest's metadata and a validation
verdict lifted into the contract's typed `summary` slot. `verify_manifest()`
recomputes the payload hash, so a consumer detects a manifest whose payload
was modified after emission. The hash recipe itself is tested against an
independently authored sibling implementation (`optimix`) in the package's
own test suite, which this vignette does not repeat.

```{r orchestra-manifest}
om <- as_orchestra_manifest(v)
om
```

`as_pesto_manifest()` bridges into `PESTO`'s narrower C2 ensemble contract
(`pesto_ensemble_manifest`), which carries a parameter ensemble alongside
the row-aligned simulated observations that ensemble produced. Exactly one
apsimR result is an object of that kind -- the posterior ensemble from
`apsim_calibrate(backend = "ies")`, which `PESTO`'s own ensemble smoother
produced. Every other result abstains with a typed reason naming the
requirement it does not meet and pointing back at
`as_orchestra_manifest()`, rather than being forced into slots it has no
basis to fill.

## Read

```{r read-prep, echo = FALSE}
sim_read_text <- if (has_apsim) {
  sprintf(
    paste0("APSIM Next Generation %s is installed, so every simulator-",
          "dependent section above ran for real. Reading the shipped ",
          "Wheat example finds %d simulation to run; predicting over it ",
          "returns %d simulated report row(s) across %d column(s)."),
    apsim_version(), length(apsim_simulations(sim)),
    nrow(pred_m@outputs), ncol(pred_m@outputs))
} else {
  paste0(
    "APSIM Next Generation is not installed on this build, so every ",
    "simulator-dependent section above took the abstention path instead ",
    "of a real run -- the same path the deliberately-broken configuration ",
    "demonstrates directly."
  )
}
```

```{r read-text, echo = FALSE, results = "asis"}
cat(sim_read_text, "\n")
```

`apsim_available()` reports `r has_apsim` on the machine that rendered this
copy of the vignette, and the deliberately-misconfigured call above abstains
with reason `r res$reason`, exactly the reason a genuinely APSIM-less build
would report from every simulator-dependent verb. The fitted Mitscherlich
curve on the synthetic N-response demonstration recovers an asymptote of
`r round(est_params$ymax, 2)`, a curvature of `r round(est_params$rate, 4)`
and a zero-input intercept of `r round(est_params$y0, 2)`, against the
generating values `4.2`, `0.018` and `1.1` used to build the synthetic data
above -- close enough, on this one draw, to show the fit recovers the
generating mechanism rather than merely interpolating it. The Latin
hypercube design table spans both parameters' full ranges (population 50 to
200, fertiliser amount 0 to 200) at its five requested points, which is what
a space-filling design over a box is meant to do at even a handful of
points.

On the twelve-point leave-one-out demonstration, every held-out point sits
close to its own leave-one-out prediction (RMSE `r round(loo_rmse, 3)`), and
the shaded 95% interval covers the truth at `r sum(covered)` of
`r length(covered)` points, exactly as the figure's colour split shows.

The validation demonstration grades a five-point synthetic pair as
`r v@metadata$verdict` (Nash-Sutcliffe efficiency `r round(v@params$nse, 2)`,
RMSE `r round(v@params$rmse, 3)`, percent bias `r round(v@params$pbias, 1)`),
and the figure shows why: the points sit close to the 1:1 line on both
sides of it rather than uniformly above or below, consistent with a percent
bias close to zero.

Lifting that validation result through `as_orchestra_manifest()` produces a
manifest whose integrity check passes (`verify_manifest(om)$ok` is
`r verify_manifest(om)$ok`) and whose headline verdict,
`r om@summary$headline`, matches the validation's own typed grade above --
the contract's typed `summary` slot is not a re-derivation, it is the same
verdict carried into a shape every other orchestra member can read without
knowing apsimR's own vocabulary.

Taken together, the answer to the opening question is yes on all three
counts: apsimR drives APSIM from R, returns results with enough provenance
(a hashed, versioned manifest) for a downstream member to trust, and, on a
build without the simulator, every one of those same verbs abstains with a
typed reason rather than failing outright, as the abstention demonstration
above shows directly.

## Limits

This vignette runs one real simulator call, if APSIM is installed, and
otherwise demonstrates the abstention path a real run would fall back on;
the calibration, sensitivity, full emulation and ground-truth sections show
only the shape of a call whose simulator cost (from dozens to several
hundred APSIM invocations) does not fit this vignette's render-time budget,
even on a build where the simulator is present -- each function's own help
page carries a runnable, simulator-gated example instead. The twelve-point
leave-one-out demonstration is a small synthetic illustration, not a
calibration certificate: coverage measured on twelve points is not a
meaningful diagnostic on its own, and the text above says so explicitly
before quoting a number from it. The Moriasi et al. (2007) verdict bands
were derived for monthly watershed and streamflow simulation, not
paddock-scale crop yield, and the typed grade in `metadata$verdict` reads
Nash-Sutcliffe efficiency alone; the RMSE and percent bias reported in the
same table are worth reading alongside it rather than trusting the single
label. The Mitscherlich (1909) attribution for the saturating dose-response
form, and the Rasmussen and Williams (2006) attribution for the exact-GP
formulation, are both cited from the literature record; neither citation was
checked against its original source in this pass, so the historical
attribution should be read as indicative rather than bibliographically
confirmed. The ground-truth demonstration illustrates the mechanism -- a
confounded assignment biases a naive comparison away from the recorded
average treatment effect -- without itself running; grading a causal method
against apsimR's OSSE output, the validation gate the Do section describes,
is not exercised here. The exact-GP backend's anisotropic fit has been
checked elsewhere against a reference implementation (DiceKriging), a
comparison this vignette does not repeat.

## What to read next

The package's function reference lists every verb this vignette does not
have room to run for real, each documented with its own runnable,
simulator-gated example. *Getting Started with PESTO* introduces the
iterative ensemble smoother that `apsim_calibrate(backend = "ies")`
delegates to. *Getting started with decideR* walks the decision layer that
reads the `orchestra_manifest` this vignette's final section builds.

## Reproduce

Seed `1` for the synthetic N-response, the Latin-hypercube design, the
leave-one-out demonstration and the calibration/sensitivity/emulation call
shapes shown above. Package versions and the APSIM version this copy was
rendered against follow.

```{r session-info}
apsim_version()
sessionInfo()
```
