---
title: "Getting started with apsimR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with apsimR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(apsimR)
has_apsim <- apsimR::apsim_available()
```

apsimR makes APSIM Next Generation -- a process-based agricultural systems
simulator -- a typed, contract-emitting member of an R analysis stack. APSIM
itself is an external .NET program; apsimR locates it, drives it, and returns its
output as a provenance-complete object that downstream calibration, uncertainty
and causal tools can consume.

## Is the simulator available?

apsimR is *standalone-functional*: it installs and loads with no simulator
present, and every simulator-dependent verb **abstains** (returns a typed reason)
rather than erroring when APSIM is absent.

```{r availability}
apsim_available()
apsim_version()
```

When the simulator is missing, a verb returns an `apsim_abstention` you can branch
on:

```{r abstention}
# A minimal simulation file, and a deliberately wrong simulator path.
f <- tempfile(fileext = ".apsimx")
writeLines(paste0('{"$type":"Models.Core.Simulations, Models",',
                  '"Name":"Simulations","Children":[]}'), f)
apsim_configure(models = file.path(tempdir(), "no-such-Models"))

res <- apsim_predict(f)
is_apsim_abstention(res)
res$reason

apsim_configure() # revert to auto-discovery
```

## Reading a simulation file

An `.apsimx` file is a typed tree of model nodes. `apsim_sim()` reads it; the
object round-trips through `apsim_write()`.

```{r sim, eval = has_apsim}
sim <- apsim_sim(apsim_example("Wheat"))
sim
apsim_simulations(sim)
```

## Prediction

`apsim_predict()` runs the simulator and returns the simulated reports as a
`predictions` manifest -- a contract-shaped object carrying the outputs plus
provenance (the APSIM version, an integrity hash).

```{r predict, eval = has_apsim}
m <- apsim_predict(apsim_example("Wheat"))
m
```

## Estimation

`apsim_estimate()` runs the simulator across an input and fits a reduced-form
mechanism (a saturating Mitscherlich dose-response by default), returning the
fitted parameters as a `parameters` manifest -- the interpretable summary a
causal or decision member consumes. The default form is the classical
saturating N-response curve of Mitscherlich (1909) `[unverified]`,

$$y(x) = y_0 + y_{\max}\left(1 - e^{-r x}\right),$$

with asymptote $y_{\max}$, zero-input intercept $y_0$ and curvature $r$
governing how quickly the response saturates. `apsim_estimate()` fits this
form by nonlinear least squares through the internal helper it calls,
`apsimR:::.apsim_fit_response()`, which is what runs underneath every
`apsim_estimate(sim, x, y)` call once the simulator is available:

```{r estimate-fit}
set.seed(1)
n_rate <- seq(0, 200, by = 10)
yield  <- 1.1 + 4.2 * (1 - exp(-0.018 * n_rate)) +
  stats::rnorm(length(n_rate), 0, 0.1)
params <- apsimR:::.apsim_fit_response(n_rate, yield, response = "mitscherlich")
m <- apsim_manifest("parameters", "demo", params = params)
m@params
```

## The forward model: one machine for three verbs

Calibration, sensitivity analysis and emulation are the same operation seen three
ways -- vary some APSIM parameters, run the simulator, reduce each run to a small
observation vector. `apsim_forward_model()` builds that operation once, as a
closure of the orchestra-standard shape `function(theta) -> obs`: `theta` is an
`nreal x npar` matrix (one row per parameter set), `obs` an `nreal x nobs` matrix.
A single APSIM invocation per row applies the edits and runs the file in one
`--apply` config; failed runs become `NA` rows rather than aborting the sweep.

```{r forward, eval = FALSE}
fm <- apsim_forward_model(
  apsim_example("Wheat"),
  parm_paths = "[Fertilise at sowing].Script.Amount",
  output = "Yield")
fm(matrix(c(0, 100, 200), ncol = 1L)) # three nitrogen rates -> three yields
```

`apsim_design()` builds the parameter points a sweep evaluates -- a space-filling
Latin hypercube by default. It is pure (no simulator), so it runs anywhere:

```{r design}
apsim_design(c("[Sow].Script.Population", "[Fertilise].Amount"),
             lower = c(50, 0), upper = c(200, 200), n = 5L, seed = 1L)
```

## Calibration

`apsim_calibrate()` solves the inverse problem: find the parameters whose
simulated observations best match data, over a box. The `optim` backend returns a
bounded least-squares point estimate with no extra dependency; the `ies` backend
(default when `PESTO` is installed) returns a posterior parameter ensemble through
an iterative ensemble smoother, with RTPS inflation guarding against ensemble
collapse. Several targets calibrate jointly through one weighted misfit.

```{r calibrate, eval = FALSE}
apsim_calibrate(
  apsim_example("Wheat"),
  parm_paths = "[Fertilise at sowing].Script.Amount",
  lower = 0, upper = 250,
  observed = c(Yield = 2200), output = "Yield",
  backend = "ies", n_real = 50L, seed = 1L)
```

## Sensitivity analysis

`apsim_sensitivity()` asks which inputs drive an output. Morris screening (the
default) ranks factors by mean absolute elementary effect (`mu_star`) at a cost of
r(p + 1) runs and flags interaction or non-linearity (`sigma`); the Sobol-Jansen
method decomposes the output variance into first-order (`S`) and total (`T`)
indices at a higher cost of n(p + 2) runs. Both use the `sensitivity` package over
the same forward model.

```{r sensitivity, eval = FALSE}
apsim_sensitivity(
  apsim_example("Wheat"),
  parm_paths = c("[Fertilise at sowing].Script.Amount",
                 "[Sow using a variable rule].Script.Population"),
  lower = c(0, 50), upper = c(250, 200),
  output = "Yield", method = "morris", morris_r = 10L, seed = 1L)
```

## Emulation

A single APSIM run costs seconds; a sweep needs thousands. `apsim_emulate()` runs
the simulator at a space-filling design once, fits a Gaussian-process surrogate,
then predicts anywhere in microseconds with a calibrated uncertainty. Every
emulator is returned with mandatory leave-one-out diagnostics -- an emulator whose
honesty is unmeasured is worse than none. The default backend is a genuine exact
GP with an automatic-relevance-determination (ARD) squared-exponential kernel,

$$k(x, x') = \sigma_f^2 \exp\!\left(-\tfrac{1}{2}\sum_{j=1}^{p}
  \frac{(x_j - x_j')^2}{\ell_j^2}\right) + \sigma_n^2 \,\mathbb{1}[x = x'],$$

with a per-dimension length-scale $\ell_j$, signal variance $\sigma_f^2$ and
noise variance $\sigma_n^2$ all fitted by maximising the exact log marginal
likelihood (Rasmussen and Williams, 2006, *Gaussian Processes for Machine
Learning* `[unverified]`), not fixed by a heuristic. It needs no extra package
and is exercised directly here (via the internal training/prediction pair
`apsimR:::.apsim_gp_train()` / `apsimR:::.apsim_gp_predict()` that
`apsim_emulate()` itself calls) on a one-dimensional test function, with a
genuine leave-one-out pass -- each point predicted from a GP refit on the other
eleven -- rather than a stand-in smoother:

```{r emulate-demo, fig.cap = "Leave-one-out predictions from the exact GP against the true test function, with the fitted 95% interval. Points falling outside their own interval are prediction failures the LOO pass is designed to catch; the caption below reports how many did."}
set.seed(1)
truth <- function(x) sin(x) + 0.1 * x
xtr <- matrix(seq(0, 10, length.out = 12), ncol = 1L)
ytr <- truth(xtr[, 1L])

# A genuine leave-one-out pass: refit the GP on the other 11 points and
# predict the held-out one, for each point in turn.
loo <- t(vapply(seq_len(nrow(xtr)), function(i) {
  gp <- apsimR:::.apsim_gp_train(xtr[-i, , drop = FALSE], ytr[-i])
  pr <- apsimR:::.apsim_gp_predict(gp, xtr[i, , drop = FALSE])
  c(mean = pr$mean, sd = pr$sd)
}, numeric(2)))
loo_rmse <- sqrt(mean((ytr - loo[, "mean"])^2))
lo <- loo[, "mean"] - 1.96 * loo[, "sd"]
hi <- loo[, "mean"] + 1.96 * loo[, "sd"]
covered <- ytr >= lo & ytr <= hi

plot(xtr[, 1L], ytr, pch = 16, xlab = "x", ylab = "y",
     ylim = range(c(lo, hi, ytr)),
     main = sprintf("Exact-GP LOO: RMSE = %.3f, 95%% coverage = %d / %d",
                     loo_rmse, sum(covered), length(covered)))
lines(xtr[, 1L], loo[, "mean"], col = "steelblue")
segments(xtr[, 1L], lo, xtr[, 1L], hi, col = "steelblue")
```

Every held-out point sits close to its own leave-one-out prediction (RMSE
`r round(loo_rmse, 3)` on this design), and the vertical bars -- the fitted 95%
interval at each held-out point -- cover the truth at `r sum(covered)` of
`r length(covered)` points here. A single twelve-point draw is far too small
to certify calibration either way; the reported `coverage95` from a full
`apsim_emulate()` run (typically evaluated over dozens of held-out points
against a design of thirty or more) is the number to trust, and it can --
and, on at least one independent Branin-function check, did -- run below its
nominal 95%, so read it rather than assume the interval is calibrated.

On the simulator the same idea fits an emulator of yield against nitrogen, with
`apsim_emulate()` reporting the leave-one-out RMSE, R-squared and 95% coverage,
and `predict()` emulating new rates:

```{r emulate, eval = FALSE}
em <- apsim_emulate(
  apsim_example("Wheat"),
  parm_paths = "[Fertilise at sowing].Script.Amount",
  lower = 0, upper = 250, output = "Yield", n = 30L, seed = 1L)
em                                       # prints the LOO diagnostics
predict(em, matrix(c(75, 175), ncol = 1L))
```

## Validation

`apsim_validate()` confronts simulated output with observations and reports the
standard agronomic goodness-of-fit metrics -- error magnitude (RMSE, MAE),
direction (mean error, percent bias) and skill relative to the observed mean
(Nash-Sutcliffe efficiency, Willmott's index of agreement, RSR) -- with a typed
verdict against the performance bands of Moriasi et al. (2007). It accepts an
`apsim_predict()` manifest, a data frame (optionally joined to the observations on
a key such as date), or plain vectors:

```{r validate}
observed  <- c(2.1, 3.4, 4.0, 5.2, 6.1)
simulated <- c(2.0, 3.6, 3.8, 5.1, 5.9)
v <- apsim_validate(simulated, observed)
v@params           # the metrics
v@metadata$verdict # the Moriasi grade
```

An observed-versus-predicted plot is available through `apsim_validation_plot()`
when `ggplot2` is installed:

```{r validate-plot, eval = requireNamespace("ggplot2", quietly = TRUE), fig.cap = "Simulated against observed values for the five-point demonstration pair, with the 1:1 line of perfect agreement. Points close to the line and on both sides of it (rather than all above or all below) indicate no systematic bias, consistent with the near-zero percent bias reported in `v@params`."}
apsim_validation_plot(v)
```

## Ground truth: APSIM as a causal test-bench

A real experiment never observes both potential outcomes of the same unit, so the
true causal effect is unknowable and a causal method can only be trusted, never
scored. A process simulator escapes that: run the same unit under control and under
treatment and both outcomes are known. `apsim_ground_truth()` does this across
heterogeneous units to manufacture an observing-system simulation experiment
(OSSE) -- a dataset of observed `(covariates, treatment, outcome)` whose true
individual and average treatment effects are recorded alongside. Treatment can be
assigned at random, or confounded by a covariate that also moves the outcome, so a
causal member (such as the orchestra's TACI / kernR tools) can be handed the
observed data and graded against the truth it cannot see.

```{r ground-truth, eval = FALSE}
units <- data.frame(
  `[Sow using a variable rule].Script.Population` = c(80, 120, 160, 200),
  check.names = FALSE)
gt <- apsim_ground_truth(
  apsim_example("Wheat"),
  treatment = "[Fertilise at sowing].Script.Amount",
  control = 0, treated = 150, units = units, output = "Yield",
  assignment = "confounded",
  confounder = "[Sow using a variable rule].Script.Population", seed = 1L)

gt                   # prints the true ATE and the (biased) naive difference
head(gt@observed)    # (covariates, W, Y) -- the input to a causal method
gt@truth$ate         # the answer key
```

Under confounding the naive treatment-minus-control difference is biased away from
`gt@truth$ate`; a causal method that conditions on the covariates should recover
the recorded ATE. That recovery is the validation gate for the orchestra's causal
members.

## Composing with the orchestra

Every result-returning verb produces an `apsim_manifest`, a versioned,
provenance-complete object of its own. A bridge to `PESTO`'s ensemble-manifest
contract, `as_pesto_manifest()`, exists for when `PESTO` is installed, but it is
not yet a working seam: `apsim_manifest` currently writes `method` values (for
example `"apsim:predict"`) that `PESTO`'s validator does not recognise, so the
bridge abstains rather than converts. Until that is closed, treat `apsim_manifest`
objects as apsimR's own typed result, not as something that already flows
directly into the calibration, uncertainty and causal stack.
